Master thesis

### 2017-05-31 docs/master-thesis.tex
03-Model v1

\setlength{\droptitle}{-2em}
  \title{Master thesis}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Sylvain SCHMITT}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2017-05-30}

@@

\caption{\label{tab:traits}Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.}
\centering
\begin{tabular}[t]{l|l|l}
\hline
Abbreviation & Description & Units\\
\hline
\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}-2\$\\
\hline
\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}-1\$\\
\hline
\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}-1\$\\
\hline
\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}-3\$\\
\hline
\$dbh\_\{thresh\}\$ & diameter at breasth height threshold & \$m\$\\
\hline
\$h\_\{lim\}\$ & asymptotic height & \$m\$\\
\hline
\$a\_h\$ & parameter of the tree-height-dbh allometry & \$m\$\\
\hline
\end{tabular}
\end{table}

@@

A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
model the abiotic environment. For each tree crown, leaf area density is
calculated on tree geometry assuming a uniform distriution across voxels
occupied by the crown. Leaf area density is computed within each voxel
summing all tree crowns inside the voxel \(v\), and is denoted
\(LAD(v)\) (leaf area per voxel in \(m².m^-3\)). The vertical sum of
\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
area per fround area in \(m^2.m^-2\) commonly called leaf area index):

@@

Daily variations in light intensity (photosynthetic photon flux density
PPFD in \(\mu mol_{photons}.m^-2.s^-1\)), temperature (T in degrees
Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
assess carbon assimilation within each voxel of the canopy and for a
representative day per month (see Appendix 1 from \citet{Li} for further
details). Variation of PPFD Within the canopy is calculated as a loacal
Beer-Lambert extinction law:

@@

The daily maximum incident PPFD at the top of canopy
\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
assumed as constant (see table \ref{tab:par} for parameters value),
besides is variation with zenith angle and species leaf inclination
angle \citep{Meir2000}. Moreover only vertical light diffusion is
considered ignoring lateral light diffusion, which can have an important
role especially in logging gaps. Finally, intra-day variation at half
hour time steps \(t\) for a representative day every month are used to
compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
trees are not simulated.

@@

@@

\subsection{Carbon uptake}\label{carbon-uptake}

@@

@@

@@

>>>>

\setlength{\droptitle}{-2em}
  \title{Master thesis}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Sylvain SCHMITT}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2017-05-31}

@@

\caption{\label{tab:traits}Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.}
\centering
\begin{tabular}[t]{l|l|l}
\hline
Abbreviation & Description & Units\\
\hline
\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}\{-2\}\$\\
\hline
\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
\hline
\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
\hline
\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}\{-3\}\$\\
\hline
\$dbh\_\{thresh\}\$ & diameter at breasth height threshold & \$m\$\\
\hline
\$h\_\{lim\}\$ & asymptotic height & \$m\$\\
\hline
\$a\_h\$ & parameter of the tree-height-dbh allometry & \$m\$\\
\hline
\end{tabular}
\end{table}

@@

A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
model the abiotic environment. For each tree crown, leaf area density is
calculated on tree geometry assuming a uniform distriution across voxels
occupied by the crown. Leaf area density is computed within each voxel
summing all tree crowns inside the voxel \(v\), and is denoted
\(LAD(v)\) (leaf area per voxel in \(m².m^{-3}\)). The vertical sum of
\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
area per fround area in \(m^2.m^{-2}\) commonly called leaf area index):

@@

Daily variations in light intensity (photosynthetic photon flux density
PPFD in \(\mu mol_{photons}.m^{-2}.s^{-1}\)), temperature (T in degrees
Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
assess carbon assimilation within each voxel of the canopy and for a
representative day per month (see Appendix 1 from \citet{Li} for further
details). Variation of PPFD Within the canopy is calculated as a loacal
Beer-Lambert extinction law:

@@

The daily maximum incident PPFD at the top of canopy
\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
assumed as constant, besides is variation with zenith angle and species
leaf inclination angle \citep{Meir2000}. Moreover only vertical light
diffusion is considered ignoring lateral light diffusion, which can have
an important role especially in logging gaps. Finally, intra-day
variation at half hour time steps \(t\) for a representative day every
month are used to compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
trees are not simulated.

@@

\subsubsection{Theory}\label{theory}

Troll simulates the carbon uptake of each individual with the Farquhar,
von Caemmerer and Berry model of C3 photosynthesis \citep{Farquhar1980}.
Gross carbon assimilation rate (\(A\) in
\(\mu mol~CO_2. m^{-2}.s^{-1}\)) will be the minimum of eiter Rubisco
activity (\(A_v\)) or RuBP generation (\(A_j\)):

\begin{equation}
  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
  \label{eq:A}
\end{equation}

\(V_{cmax}\) is the maximum rate of carboxylation
(\(\mu mol~CO_2.m^{-2}.s^{-1}\)). \(c_i\) is the \(CO_2\) partial
pressure at carboxylation sites. \(\Gamma^*\) is the \(CO_2\)
compensation point in absence of dark respiration. \(K_m\) is the
apparent knietic constant of the Rubisco. And \(J\) is the electron
transport rate (\(\mu mol e^-.m^{-2}.s^{-1}\)). \(J\) depends on the
light intensity with \(PPFD\):

\begin{equation}
  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
  \label{eq:J}
\end{equation}

\(J_{max}\) is the maximal electron transport capacity
(\(\mu mol e^-.m^{-2}.s^{-1}\)). \(\theta\) is the curvature factor. And
\(\alpha\) is the apparent quantum yield to electron transport
(\(mole^-.mol~photons^{-1}\)).

Carbon assimilation by photosynthesis will then be limited by the
\(CO_2\) partial pressure at carboxylation sites. Stomata controls the
gas concentration at carboxylation sites throught stomatal transport:

\begin{equation}
  A = g_s*(c_a-c_i)
  \label{eq:Ag}
\end{equation}

\(g_s\) is the stomatal conductance to \(CO_2\)
(\(molCO_2.m^{-2}.s^{-1}\)). TROLL simulates stomatal conductance
\(g_s\) with the model from \citep{Medlyn2011}:

\begin{equation}
  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
  \label{eq:gs}
\end{equation}

\(g_0\) and \(g_1\) are parameters from the model. TROLL model assume
\(g_0 \approx 0\) (empirically tested and considered as reasonable).

\subsubsection{Parametrization}\label{parametrization}

Leaf traits can be used as proxy of photosynthesis, especially leaf
nutrient content which directly play a role in it
\citep{wright_worldwide_2004}. \citet{Domingues2010} suggested that
\(V_{cmac}\) and \(J_{max}\) were both limited by the leaf concentration
of nitrogen \(N\) and phosphorus \(P\) (\(mg.g^{-1}\)):

\begin{equation}
  log_{10} V_{cmax-M} = min( 
  \begin{array}{c} 
    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
  \end{array} 
  )
  \label{eq:VcmaxM}
\end{equation}

\begin{equation}
  log_{10} J_{max-M} = min(
  \begin{array}{c} 
    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
  \end{array}
  )
  \label{eq:JmaxM}
\end{equation}

\(V_{cmax-M}\) and \(J_{max-M}\) are the photosynthetic capacities at
\(25^\circ C\) of mature leaves per leaf dry mass (resp.
\(\mu mol CO_2.g^-1.s^{-1}\) and \(\mu mol e^-.g^{-1}.s^{-1}\)). \(LMA\)
is the leaf mass per are (\(g.cm^{-2}\)). \(V_{cmax}\) and \(J_{max}\)
are calculated by multiplying \(V_{cmax-M}\) and \(J_{max-M}\) by
\(LMA\). \(V_{cmax}\) and \(J_{max}\) variation with temperature are
caluclated with \citet{Bernacchi2003} (see Appendix 2 from \citet{Li}
for further details).

TROLL computes leaf carbon assimilation \(A_l\) combining equations from
\eqref{eq:A} to \eqref{eq:JmaxM} for each tree crown voxel within in each
crown layer \(l\):

\begin{equation}
  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
  \label{eq:Al}
\end{equation}

\(PPFD_{month}(v,t)\), \(VPD_{month}(v,t)\) , and \(T_{month}(v,t)\) are
derived from microclimatic data. \(n_v\) is the number of voxels within
crown layer \(l\). And the sum is calculated over the \(t_M\)
half-hourly intervals \(t\) of a tipical day.

@@

A large fraction of plants carbon uptake is actually used for plant
maintenance and growth respiration. The autotrophic respiration can
represents up to 65\% of the gross primary productivity but varies
strongly among species, sites, and environnements.

TROLL uses \citet{Atkin2015} database of mature leaf dark respiration
and associated leaf traits to compute leaf maintenance respiration:

\begin{equation}
  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
  \label{eq:Rl}
\end{equation}

\(R_{leaf-M}\) si the dark respiration rate per leaf dry mass at a
temperaure of \(25^\circ C\) (\(nmolCO_2.g^{-1}.s^{-1}\)). The other
terms are in equations \eqref{eq:VcmaxM} and \eqref{eq:JmaxM}. TROLL assume
leaf respiration during day light to be 40\% of leaf dark respiration,
and computes total leaf respiration by accounting for the legnth of
daylight.

TROLL model stem respiration (\(R_{stem}\) in \(\mu molC.s^{-1}\)) with
a constant respiration rate per volume of sapwood:

\begin{equation}
  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
  \label{eq:Rs}
\end{equation}

dbh, h, CD and ST are tree diameter at breast height, height, corwn
depth and sapwoond thickness, respectively (\(m\)). TROLL assumes
\(ST=0.04~m\) when \(dbh>30~cm\) and an increasing \(ST\) for lower
\(dbh\).

Finally, TROLL computes both fine root maintenance respiration, as half
of the leaf maintenance respiration. Whereas coarse root and branch
maintenance respiration is computed as half of the stem respiration. And
growth respiration (\(R_{growth}\)) is assumed to account for 25\% of
the gross primary productivity minus the sum of maintenance
respirations.

\subsection{Net carbon uptake}\label{net-carbon-uptake}

Net primary production of carbon for one individual \(NPP_{ind}\)
(\(gC\)) is computed by the balance between gross primary production
\(GPP_{ind}\) and respirations \(R\):

\begin{equation}
  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
  \label{eq:NPP}
\end{equation}

TROLL partitions individuals total leaf area \(LA\) into three pools for
different leaf age classes corresponding to different photosynthesis
efficiency (young, mature and old leaves with \(LA_{young}\),
\(LA_{mature}\), and \(LA_{old}\) respectivelly). Consequently we can
compute growth primary production for one individual as:

\begin{equation}
  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
  \label{eq:GPP}
\end{equation}

h and CD are tree height and crown depth, repectivelly (\(m\)).
\(\lfloor x \rfloor\) is the rounding function. \(\Delta t\) is the
duration of a timestep (\(year\)).

Thus, TROLL can compute carbon allocation to wood into an increment of
stem volume \(\Delta V\) (\(m^3\)):

\begin{equation}
  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
  \label{eq:DeltaV}
\end{equation}

\(f_{wood}\) is the fixed fraction of NPP allocated to stem and
branches. \(wsg\) is the wood specific gravity (\(g.cm^{-3}\), see
\ref{tab:traits}). TROLL assume large trees less efficient to convert
NPP as growth by using a size-related growth decline with function
\(Senesc\) after a specific diameter at brest height threshold
\(dbh_{thresh}\):

\begin{equation}
  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
  \label{eq:Senesc}
\end{equation}

Finally, TROLL can compute carbon allocation to canopy with canopy NPP
fraction denoted \(f_{canopy}\) and decomposed into leaf, twig and fruit
production. Carbon allocation to leaf results in a new young leaf pool,
whereas other leaf pools are updated as follow:

\begin{equation}
  \begin{array}{c} \\
   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
  \end{array}
  \label{eq:DeltaLA}
\end{equation}

\(\tau_{young}\), \(\tau_{mature}\), and \(\tau_{old}\) are species
residence times in each leaf pools (\(years\)). The sum of residency
time thus defined the leaf lifespan
\(LL = \tau_{young} + \tau_{mature} + \tau_{old}\) (\(years\)).
\(\tau_{young}\) is set to one month and \(\tau_{mature}\) is set to a
third of leaf lifespan \(LL\). Previous implementation of TROLL model
used \citet{Reich1991a} allometry to infer leaf lifespan \(LL\) from
species leaf mass per area \(LMA\) \citep{Li}. But the use of the
allometrie from \citet{Reich1991a} with current implementation of the
TROLL model resulted in an underestimation of leaf lifespan for low LMA
species. Consequently in the following paragraph we suggest a new
allometry. Belowground carbon allocation is not simulated inside TROLL.

\subsection{Leaf lifespan}\label{leaf-lifespan}

The underestimation of leaf lifespan for low LMA species with the
allometry from \citet{Reich1991a} resulted in indivduals unealistic
early death from carbon starvation. We gathered data from
TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
\citep{wright_worldwide_2004} datasets. We used an out of the bag method
applied on a random forest to select variables with highest importance
to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
used a bayesian approach to test different models with growing level of
complexity. The model with the best tradeoff between complexity (number
of parameters), convergence, likelihood, and prediction quality (root
mean square error of prediction RMSEP) was kept. We selected following
model with a maximum likelihood of 13.6 and a RMSEP of 12 months:

\begin{equation}
  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
  \label{eq:LLb}
\end{equation}

Leaf lifespan \(LL\) follows a lognormal law with location infered from
leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
presemption following a gamma law. \(d\) represents the dataset random
effects and encompass environmental and protocol variations. The
sampling of model \eqref{eq:LLb} resulted in the following allometry:

\begin{equation}
  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
  \label{eq:LL}
\end{equation}

@@

Once the increment of stem volume \(\Delta V\) calculated with equation
\eqref{eq:DeltaV}, TROLL convert it into an increment of tree diameter at
breast height denoted \(\Delta dbh\). TROLL infer tree height from
\(dbh\) using a Michaelis-Menten equation:

\begin{equation}
  h = h_{lim}*\frac{dbh}{dbh + a_h}
  \label{eq:h}
\end{equation}

On the other hand, we have the trunk volume
\(V = C * \pi * (\frac{dbh}{2})^2*h\), thus:

\begin{equation}
  \begin{array}{c} \\
    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
  \end{array}
  \label{eq:Deltadbh}
\end{equation}

Next, TROLL used the new trunk dimension (\(dbh\) and \(h\)) to update
tree crown geometry using allometric equations \citep{Chave2005}:

\begin{equation}
  \begin{array}{c} \\
    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
  \end{array}
  \label{eq:C}
\end{equation}

Finally, TROLL computes the mean leaf density within the crown (\(LD\)
in \(m^2.m^{-3}\)) assuming a uniform distribution:

\begin{equation}
  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
  \label{eq:LD}
\end{equation}

Mortality is partitioned in three factors inside TROLL: background death
\(d_b\), treefall death \(d_t\) and negative density dependent death
\(d_{NDD}\). Because density dependent death \(d_{NDD}\) is still in
development inside TROLL we did not used it, so we will not detail is
computation.

\citet{chave_towards_2009} advocated for a wood economics spectrum
opposing fast growing light wood species species with high risk of
mortality to slow growing dense wood species with reduced risk of
mortality. Hence, background mortality is derived from wood specific
gravity \(wsg\) inside TROLL:

\begin{equation}
  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
  \label{eq:db}
\end{equation}

\(m\) (\(events.year^{-1}\)) is the reference background death rate for
lighter wood species (pioneers). \(d_n\) represents death by
carbohydrates shortage. If the number of consecutive day with
\(NPP_{ind} < 0\) \eqref{eq:NPP} is superior to tree leaf lifespan \(d_n\)
is set to 1 and remains null in other cases.

Mortality by treefall inside TROLL depends on a specifric stochastic
threshold \(\theta\):

\begin{equation}
  \theta = h_{max}*(1-v_T*|\zeta|)
  \label{eq:theta}
\end{equation}

\(h_{max}\) is the maximal tree height. \(v_T\) is the variance term set
to 0.3. \(|\zeta|\) is the absolute value of a random centered and
scaled Gaussian. If the tree hieght \(h\) is superior to \(\theta\) then
the tree may fall with a probability \(1-\theta/h\) \citep{Chave1999}.
The treefall direction is random (drawn from a uniform law
(\(\mathcal{U}[0,2\pi]\)). All tree in the trajectory of the falling
tree will be hurted through a variable denoted \(hurt\), incremented by
fallen tree height \(h\). If a tree height is inferior than its \(hurt\)
values then it may die with a probability
\(1-\frac{1}{2}\frac{h}{hurt}\). \(hurt\) variable is reset to null at
each timestep (\(month\)).

@@

Once the tree became fertile they will start to disperse seeds. TROLL
consider tree as fertile after a specific height threshold
\(h_{mature}\) \citep{Wright2005}:

\begin{equation}
  h_{mature} = -11.47+0.90*h_{max}
  \label{eq:hmature}
\end{equation}

But TROLL is not considering seed directly through a seedbank, instead
seed might be interpreted as a seedling recruitment opportunity. The
number of reproduction opportunities per mature tree is denoted \(n_s\)
and set to 10 for all species. This assumption originates from a
trade-off between seed number and seed size resulting in equivalent
survival and recruitment probability. All \(n_s\) events are dispersed
with a distance randomly drawn from a Gaussian distribution.
Additionally, TROLL model consider external seedrain through \(n_{ext}\)
events of seed immigration:

\begin{equation}
  n_{ext} = N_{tot}*f_{reg}*n_{ha}
  \label{eq:next}
\end{equation}

N\_\{tot\} is the external seedrain per hectare (number of reproduction
opportunities). \(f_{reg}\) is the species regional frequency.
\(n_{ha}\) is the simulated plot size in \(ha\).

Finally, a bank of seedlings to be recruited is defined for each pixel.
If the ground-level light reaches a species light compensation point
\(LCP\) the species will be recruited:

\begin{equation}
  LCP = \frac{R_{leaf}}{\phi}
  \label{eq:LCP}
\end{equation}

\(R_{leaf}\) is the leaf respiration for maintenance (see \eqref{eq:Rl}).
\(\phi\) is the quantum yield (\(\mu mol C.\mu mol~photon\)) set to
0.06. If several species reach their \(LCP\), one is picked at random.
Seedlings are recruited with following intial geometry:

\begin{equation}
  \begin{array}{c} \\
    dbh = \frac{a_h}{h_{max} - 1}\\
    h = 1~m\\
    CR = 0.5~m\\
    CD = 0.3~m\\
    LD = 0.8~m^2.^{-3}
  \end{array}
  \label{eq:C}
\end{equation}

@@

\subsection{Sensitivity}\label{sensitivity}

\subsection{Disturbance}\label{disturbance-1}

\subsection{Sylviculture}\label{sylviculture}
